---
title: "HydroData"
output: html_document
---
```{r}
library(devtools)
install_github("mikejohnson51/HydroData")
```

##Defining an AOI

The central componenet of every function in HydroData is a user defined area of interest (AOI) which can be defined in a number of ways using three core parameters: state, county, and clip_unit

(1) An AOI can be defined as a state name or abbriviation:

```{r}
 CA = getAOI(state = "California")
 TX = getAOI(state = "TX")
 
 par(mfrow = c(2,1))
 plot(CA); plot(TX)
```

(2) An AOI can be defined by a state, county pair:

```{r}
 sb = getAOI(state = "California", county = "Santa Barbara")
 
 plot(sb)
```

(3) An AOI can be defined A user supplied Spatial or Raster object:
 
```{r}
 la.met = getAOI(clip_unit = rgdal::readOGR("/LA_metro"))
 
 plot(la.met)
```
 
(3) An area defined by a location a bounding box height and width and optional origin:
      The centroid can be defined by:
      
```{r}
 nwc = getAOI(clip_unit = list("National Water Center", 10, 10))
 
 pt = getAOI(clip_unit = list(34.41, 119.85, 10, 10))
 
 goleta = getAOI(clip_unit = list("KMART near UCSB", 10, 10, "lowerleft"))
 
 berkley = getAOI(clip_unit = list("UC Berkley", 10, 10, "upperright"))
``` 

In all of the following functions state, county, and clip_unit are offered as parameters so that you can define your search for finding and getting data.

## Other Common Parameters

Much like state,county, and clip_unit are used to construct AOI defintions all HydroData functions share other commom parameters:

 - basemap:  When TRUE adds a basemap to a returned list
 - boundary: When TRUE adds a the spatial object defining the AOI to a returned list
 - save: When TRUE write the called data to a users disk in their working dircory in a HydroData folder
 
Other functions might have unique parameters such as year or resolution which will be covered in the examples however with these parameters the basics of all functions are known:

## Find Data Functions

#### USGS NWIS Gaging Stations

```{r}
# Find all USGS stations in El Paso County, Colorado

elpaso = findUSGS(state = 'CO', county = 'El Paso', basemap = T, boundary = T)

{plot(elpaso$basemap)
plot(elpaso$boundary, add = T, lwd = 5)
plot(elpaso$usgs, add = T, col = 'darkgreen', pch = 16)}

head(elpaso$usgs@data)
```

#### GCHN Stations

