---
pagetitle: HydroData
---

```{r echo = FALSE, results='hide', warning=FALSE,message=FALSE}
devtools::load_all()
```

## Finding Stations 

Finding data for an Area of Interest (AOI) often requires finding what observation stations are avialable. The `Hydrodata` **find** functions can locate stations from six major networks including:

1.  Airport Weather Stations (Open Flights)
2.  Global Historic Climatology Network (NOAA)
3.  National Water Information System (USGS)
4.  National Inventory of Dams (USACE)
5.  Snow Telemetry (SNRCS)
6.  Reservoirs (USGS)

If you have not yet learned how AOI queries are defined in `HydroData` I sugest reviewing the [AOI page](aoi.html) first. 

What follows are basic examples for quering information from thes

### Finding Airports

Airports offer a network of daily meterological observation stations. [Open flights](   ) has global metadata for 7,184 airports, the majority of which report information to WeatherUnderground. 

Here's a basic example for subseting the Open for all airports in El Paso County, Colorado:

```{r}
findAirports(state = 'CO', county = "El Paso") %>% explore()

```

Interacting with this map can show where/if airports exist within an AOI. 

By clicking on each icon you can see the name, city, and four digit ICAO identifier associated with each airport. 

Clicking of the name will direct you to the WeatherUnderground webpage assocaited with airport where.

To get data from these airport stations you can use the `getWU()` function which is covered in greater detail in the "GETTING TABULAR DATA" page. 


```{r, messages = FALSE}
ap = findAirports(state = 'CO', county = "El Paso", ids = TRUE)

ap$ids
```


### Finding Global Historic Climatology Network
