---
pagetitle: HydroData for R
---

```{r echo = FALSE, results='hide', warning=FALSE,message=FALSE}
devtools::load_all()
```

## The HydroData object

All `HydroData` **find** and **get** functions return a object of class `HydroData`. This page covers the key components of this class which, at its core, is a list of features related to the query that defined the object.

All objects range is size from 4 to 8 elements:

### Features

* AOI Description (*default*)
* Source of data (*default*)
* Project (if spatial) (*default*)
* Data (*default*)


* Basemaps (*optional*)
* AOI shapefile (*optional*)
* Fiat boundary shapefile (*optional*)
* IDs of critical station information (*optional*)

### Getting a HydroData Object

In this basic example we will query a `HydroData` object containing all USGS stations in a 100 square mile region centered on UCSB. If you have not yet learned how AOI queries are defined in `HydroData` I sugest reviewing the [AOI page](aoi.html) first. 

```{r}
usgs = findUSGS(clip_unit = list("UCSB", 10, 10), basemap = T, boundary = T, ids = T)
```

### Name 

The first element of a HydroData object is a text string in plain english describing the location of the AOI.

```{r}
usgs$name
```


### Source

The second element is a text string detailing the source of the data contained in the object.

```{r}
usgs$source
```

### Projection

The third element is the projection of the HydroData object, if it is spatial.

```{r}
usgs$proj
```

### Data

The fourth element is the actual data queried. This data can be a data.frame, sp, or raster object.

```{r}
usgs$nwis
plot(usgs$nwis)

```

### Basemaps

```{r}
plot(usgs$basemap)
```



### IDs

```{r}
usgs$ids
```


### AOI shapefile

```{r}
plot(usgs$boundary)
```

### Intersecting Fiat Boundary 

```{r}
plot(usgs$fiat)
```


### Overall 

```{r}
str(usgs, max = 1)
```


