---
pagetitle: HydroData
---

```{r echo = FALSE, results='hide', warning=FALSE,message=FALSE}
devtools::load_all()
```

## Introduction

All environmental and hydrologic research begins with data; the question of what data is available; and the challenge of gathering it. HydroData is designed to help users find, get, visualize and use a range of Earth Systems data providing functions to automatically download; interact with; and share data.

### Features

* Access to 19 data sources from a consistent query
* Helps *find* where data is available and  *get* the data you want
* Consistent data formats, projections, and documentation within all `HydroData` object
* Interactive mapping and plotting via `leaflet` and `dygraphs`
* Embed maps and data in [R Markdown](http://rmarkdown.rstudio.com/) and [Shiny](http://shiny.rstudio.com/) apps or export HTML files for easy documentation, reproducabilty, and sharing.
* Export data for analysis in other platforms like python, ArcMap or QGIS

### Installation

To install this package, run this command at your R prompt:

```{r eval=FALSE}
# to install the development version from Github, run
devtools::install_github("mikejohnson51/HydroData")
```

Once installed, you can use this package in the R console, within [R Markdown](http://rmarkdown.rstudio.com/) documents, and within [Shiny](http://shiny.rstudio.com/) applications.

### Basic Usage

You use HydroData with these basic steps:

1. Define your Area of Interest (AOI).
2. "Find" and "Get" data for that AOI using functions (e.g. `findUSGS`, `getNED`)
3. Explore spatial and time series data interactively (e.g. `explore`, `inspect`)
4. Repeat as desired.
5. Analysis data within the R or export for analysis in platform of choice

Here's a basic example for all finding USGS stream flow gages in Santa Barbara County, California:

```{r width = 120, height = 30}
library(HydroData)
findUSGS(state = 'CA', county = "Santa Barbara") %>% explore()
```

### Next Steps

Before exploring the rest of this site, I suggest checking out the [sources](sources.html) page to get a better idea of where data comes from, and the [Defining an AOI](aoi.html) page as it describes the core of a HydroData query. 

I then suggest checking out [HydroData Object](hydrodata.html) page to get a sense of the structure of all HydroData objects; the [finding data](find.html) page to get a better sense of what data stations and spatial features `HydroData` can help you find. The [getting data](get.html) page to get a better sense of what data `HydroData` can help you get and finally the [Additional features](additional.html) page to understand what  features are offered in the package.



