---
pagetitle: HydroData for R
---

```{r echo = FALSE, results='hide', warning=FALSE,message=FALSE}
devtools::load_all()
```

## Introduction

All environmental and hydrologic research begins with data; the question of what data is available; and the challenge of getting it in a format that is useful. `HydroData` is designed to help users find, get, and visualize a range of earth systems data providing functions to automatically download; interact with; and share data.
<br>

### Features

* Access to 19 data sources through a consistent query
* Helps *find* where data is available and  *get* the data you want
* Harmonious data formats, projections, and documentation within all `HydroData` objects
* Interactive mapping and plotting via `leaflet` and `dygraphs`
* Embed maps and data in [Rmarkdown](http://rmarkdown.rstudio.com/); [Shiny](http://shiny.rstudio.com/); or  HTML documents for easy documentation, reproducabilty, and sharing.
* Ability to export data for use in other platforms like Python, ArcMap or QGIS
<br>

### Installation

To install `HydroData`, run this command at your R prompt:
<br>
```{r eval=FALSE}
devtools::install_github("mikejohnson51/HydroData")
```

Once installed, you can use the package in the R console, within [R Markdown](http://rmarkdown.rstudio.com/) and  [Shiny](http://shiny.rstudio.com/) applications.
<br>

### Basic Usage

You use `HydroData` with these basic steps:

1. Define an Area of Interest (AOI).
2. "Find" and "Get" data for that AOI using functions (e.g. `findUSGS`, `getNED`)
3. Explore spatial and time series data interactively (e.g. `explore`,  `inspect`)
4. Repeat as desired.
5. Analyze data or export for external use

<br>
Here's a basic example for all finding USGS stream flow gages in Santa Barbara County, California:
<br>

```{r, message = FALSE}
library(HydroData)
findUSGS(state = 'CA', county = "Santa Barbara") %>% explore()
```

### Next Steps

Before exploring the rest of this site, I suggest checking out the [sources](sources.html) page to get a better idea of what data is available and where data comes from, and the [Defining an AOI](aoi.html) page as it describes the core of a `HydroData` query. <br>
<br>
I then suggest checking out [HydroData Object](hydrodata.html) page to become familiar with the structure of all `HydroData` objects, and then the [find Station](find.html), [get obsevation data](get_tab.html), [get vector data](get_vec.html), [get raster data](get_rast.html),  to get a better sense of what data stations and spatial features `HydroData` can help you find. 
<br>
<br>
Finally, I would suggest checking out the [vizualization page](viz.html) to get a better sense of how to explore and 
inspect `HydroData` objects interactively. 
<br>
<br>
Please feel free to use [Github issues](https://github.com/mikejohnson51/HydroData/issues) to file bug reports or feature requests or <a href="mailto:mike.johnson@geog.ucsb.edu?Subject=HydroData%20Inquiry" target="_top">email me</a> directly with questions or collaboration ideas. 
<br>

