---
pagetitle: HydroData
---

```{r echo = FALSE, results='hide', warning=FALSE,message=FALSE}
# to install the development version from Github, run

devtools::load_all()
```

## Introduction

All environmental and hydrologic research begins with data; the question of what data is available; and the challenge of gathering it. HydroData is designed to help users find, get, visualize and use a range of Earth Systems data by providing functions to automatically download; interactive with and share data.

### Features

* Access to 19 data sources from the R Console or Rstudio from a consistent query
* Helps *find* where data is avaialble and then *getting* the data you want
* Consistent data formats, projections, and documention within a `HydroData` object
* Interactive mapping and plotting via `leaflet` and `dygraphs`
* Embed maps and data in [knitr](http://yihui.name/knitr/)/[R Markdown](http://rmarkdown.rstudio.com/) documents and [Shiny](http://shiny.rstudio.com/) apps or export to browser for easy documentation, reproducabilty, and sharing.
* Export data for analysis in other platforms like python, ArcMap or QGIS

### Installation

To install this R package, run this command at your R prompt:

```{r eval=FALSE}
# to install the development version from Github, run
devtools::install_github("mikejohnson51/HydroData")
```

Once installed, you can use this package at the R console, within [R Markdown](http://rmarkdown.rstudio.com/) documents, and within [Shiny](http://shiny.rstudio.com/) applications.

### Basic Usage

You use HydroData with these basic steps:

1. Define your Area of Interet (AOI).
2. "Find" and "Get" data for that AOI using functions (e.g. `findUSGS`, `getNED`)
4. Explore spatial and timeseries data interactivly (e.g. `explore`, `inspect`)
3. Repeat as desired.
4. Use data within the R console or R studio or export for analysis in platform of choice

Here's a basic example for finding USGS streamflow gages in Santa Barbara County, California:

```{r width = 120, height = 30}
library(HydroData)
findUSGS(state = 'CA', county = "Santa Barbara") %>% explore()
```

### Next Steps

We suggest checking out to [AOI](sources.html) page before exploring the rest of this site, as it describes core of the we'll use throughout the examples on the other pages.
